---
output: github_document
bibliography: inst/REFERENCES.bib
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

library(weatherImpute)
```

# weatherImpute

<!-- badges: start -->
<!-- badges: end -->

weatherImpute is a package designed to allow easy imputation of missing observations
in daily temperature and precipitation observations. Furthermore, it allows a 
systematic comparison of different imputation techniques and an evaluation of these.
weatherImpute should help users to make a true decision about the imputation method.

## Installation

You can install the development version of weatherImpute from [GitHub](https://github.com/larscaspersen/weatherImpute) with:

``` r
# install.packages("devtools")
devtools::install_github("larscaspersen/weather_impute")
```

## Package content

There are three main functions to use from weatherImpute:

* `patch_flexible_several_stations()` to patch gaps in a set of weather stations 
* `get_eval_one_station()` to calculate imputed weather observations for existing observations
* `get_eval_metrics()` to calculate performance scores based on the output of the previous function

Furthermore, the package contains a set of imputation methods, which is by far
not exhaustive. The selection of here presented imputation methods is not a recommendation but simply a collection of functions to test the packages functionality.
The previous mentioned functions were designed in such a way, that the inclusion of user-defined imputation methods should be relatively easy. The idea of weatherImpute is more of a patching-comparison framework than a complete list of recommended methods. Currently weatherImpute contains the following patching functions, some of them are simple wrapper functions of other packages.

* `patch_amelia()`: multiple imputations using the Amelia package
* `patch_climatol()`: iterative normal ratio using the climatol package
* `patch_forest()`: imputation using random forest of the missForest package
* `patch_idw()`: imputation using inverse distance weighting
* `patch_mean()`: imputation using mean of closest neighbours observations
* `patch_mice()`: multiple imputation using mice and micemd package
* `patch_normal_ratio()`: imputation using the adjusted normal ratio method
* `patch_pca()`: imputation using probebalistic principle component analysis (ppca) or NIPALS method of the pcaMethods package

Furthermore, the package contains several evaluation scores. Similar to imputation methods there is a myriad of methods used in weather imputation studies. Almost all studies either use the root mean square error (RMSE) or the mean absolute error (MAE). But especially for precipitation these methods fall short, because not only the precipitation amount but also the occurrence is important when evaluating imputed precipitation. A summary score can be calculated as well, which gives a harmonized overall score, taking all the evaluation scores specified in `get_eval_metrics` into account, following the approach of @teegavarapu_missing_2014. 
Again, similar to the patching function, emphasis has been put on the possibility to use user-defined evaluation scores or functions of other R-packages.
Currently the following evaluation metrics are available in weatherImpute

* `calc_d_index()` refined index of agreement
* `calc_hanssen_kuipers()` score for preciptiation occurrence
* `calc_hit_score()` score for precipitation occurrence
* `calc_index_agreement()`
* `calc_KS_test()` p-value of Kolmogorov-Sminorv test which checks if two vectors come from the same continous distribtution 
* `calc_MAE()` mean absolute error
* `calc_MCC()` Matthews correlation coefficient, a score used in classification problems (like precipitation occurrence), which is said to handle imbalanced classes (rain / no rain) well
* `calc_NSE()` Nash-Sutcliffe model efficiency coefficient
* `calc_S_index()` similarity index
* `calc_skill_score()` skill score

## How to use the package

```{r}
library(weatherImpute)
## basic example code

#to impute several stations
complete_weather <- patch_flexible_several_stations(weather = weather_Tmin, 
                                target = c('cimis_2', 'cimis_15'), 
                                weather_info = weather_info, 
                                method = 'patch_normal_ratio',
                                method_patches_everything = F)

#compare imputed to actual observations
patched <- get_eval_one_station(weather = weather_Tmin,
                     weather_info = weather_info,
                     target = 'cimis_2', 
                     patch_methods = c('patch_idw','patch_normal_ratio'), 
                     method_patches_everything = c(TRUE, FALSE))

#bring result to long format
patched_long <- reshape2::melt(patched, 
                               measure.vars =c('patch_idw','patch_normal_ratio'),
                               variable.name = 'patch_method')

#calculate evaluation scores
patch_eval <- get_eval_metrics(eval_df = patched_long)


```

## How to prepare your input data

*This part is still under construction and may be incomplete.*

The input data should be organized in data.frames, one object for measured variable. weatherImpute currently supports minimum daily temperature, maximum daily temperature and daily precipitation sums. An example of the input data can be seen here

```{r }

head(weather_Tmin)

```

Columns `Date`, `Day`, `Month` and `Year` are mandatory (also the spelling of the column name needs to be the same). Every additional column is assumed to be a weather statio. Column names of weather stations need to be unique and the same as in the `id` column of second object needed, which is called `weather_info`. Here is a glimpse of its structure

```{r }

head(weather_info)

```

Important are the columns `id`, `Latitude` and `Longitude`. Coordinates need to be in the decimal format. The `Name` is less important, but some patching functions do not work, if also all entries in that column are unique.

## How to add user-defined patching and evaluation functions?

An important feature is the compatibility of weatherImpute functions and user-defined patching or evaluation functions. Here is an example how to integrate a user-defined function in `patch_flexible_several_stations`, but the concept is the same for `get_eval_one_station`. user-defined function need to be present in the work environment. The user-defined patching function needs to fulfil some general features:

* it takes `weather` as an input and `weather` needs to be organized using the same principles as in `weather_Tmin
* it takes a second object called `weather_info` as an input witht the ids and coordinates of the weather stations. id column and column names in weather need to be the same. Even if the user-defined function does not actually need the object, it is still important that the argument is present in the function call. To put it plain: it needs to be there, no matter if the patching function actually makes use of it
* the same is true for the last mandatory argument called `target`, which is a character of the weather station id. Some functions like `patch_climatol` automatically impute data for all weather stations (which makes target obsolete) but the weatherImpute functions still require `target` to be an argument. If it is not needed, then simply putt it as NULL

In most cases patching functions have more arguments. These can be later adjusted in the function call with the argument `additional_input` of `patch_flexible_several_stations` which is a named list of all further specified arguments.

```{r }


#example of user-defined patching method
#maybe built UK traditional method which is also used in chillR


#check if function works on weather_Tmin

#integrate in patch_flexible_several_stations call

```

The same can be done with user-defined evaluation functions. Also functions of other packages can be used, the example below also shows how to add the RPIQ function of the package chillR. Packages do not necissarily need to be loaded, but installed of course. The evaluation methods should follow a simple structure: they should have a `predicted` argument where the patching functions output goes, an `observed` argument where the actually observation of the weather station goes and they need to handle missing values (NAs).

```{r }


#built user defined evaluation metric

#integrate in get_eval_metrics call; also add metrics of other packages like RPIQ of chillR

```

## References